User-level Grid Monitoring with Inca 2

### 2007-02-21 pubs/papers/hpdc-gmw-2007/paper.tex
smoothing of intro and biblio adding and cleaning
git-svn-id: file:///localdisk/subversion/inca/trunk/pubs@6531 7dba3f4a-8be6-0310-8b3b-b4fec25ea7f3

% Complex means they are difficult to provide and maintain. Describe the
% stakeholders.   Monitoring is needed.
Providing and maintaining a stable infrastructure for these complex 
Grid systems poses challenges for both administrators who build and
maintain Grid resources and scientists who use them.  
% Why?  
% 1. Administration is distributed across multiple admins
Administration of the Grid infrastructure is typically coordinated by a group 
of Grid operators who decide when software should be deployed to resources and
how it should be configured.  Each site's system administrators are then
responsible for providing the required software components and services for
their resource.  Grid operators need to take into account local site policies
and heterogenous architectures when requiring software.  System administrators
need to learn enough Grid knowledge to deploy software and debug problems.  In
this distributed model, good communication of the requirements is required in
order for Grid resources to interoperate with one another.  Otherwise
inconsistencies between the resources arise that can either inconvenience the
user or prevent them from using particular resources together.
% 2. Grid software failures
distributed system, components will fail over time.  As with any distributed
systems, the probability of failure increases as the number of elements 
increase.  Failures can occur due to system downtimes, misconfiguration in the 
software, bugs in the software, high load.  
which means miscommunications can occur.  
% 3. Changes 
Third it is challenging due to the
activeness in Grids, packages and updates can occur frequently.

% What types of Grid monitoring are around?  What user-level monitoring
% provides and why we think it is good
One approach to monitoring Grids, used by tools such as
MonALISA~\cite{monalisa} and GridICE~\cite{gridice}, is to aggregate and
display data from existing cluster or system administrator monitoring tools
such as Ganglia~\cite{ganglia}, CluMon~\cite{clumon}, and
Nagios~\cite{nagios}.  This provides a centralized, systems-level view of Grid
resources where low-level host statistics and queue information can be
examined.  This type of monitoring information is useful for showing the
utilization of Grid resources, however, it does not provide the type of
high-level monitoring needed to detect user problems within the Grid
infrastructure, such as incompatible software versions.  User-level
Grid monitoring approaches testing is important because it run commands that users use and tests
to see if they operate as expected. Execute commands published in user
documentation, detect problems otherwise reported by users.

so called user-level Grid monitoring is bIn our view, what was
needed was a separate
monitoring system to perform automated, periodic user-level Grid testing with
the following features:
User-level Grid monitoring is x and is valuable.  
At that
time, the only available user-level Grid monitoring tools were the NCSA
TestGrid script [9] and GITS [10], scripts that run a fixed number of Grid
tests and output results in HTML.  Although these tools were easy to install
and produced useful information, they showed only the view of the Grid from a
single resource, lacked automation, and were not easily extensible.

% Inca 1~\cite{inca1} first addressed user-level Grid monitoring but had
% limitations.  We learned lessons from our TeraGrid deployment.  
In 2003, SDSC, in partnership with TeraGrid, began developing the Inca
monitoring tool to implement a user-level monitoring system with the above
features and started using it to validate and verify that CTSS was deployed
consistently across all TeraGrid resources and to monitor its status.  
The initial version of Inca was implemented as a client-server architecture.
It provided data collection, data storage (with limited archiving
capabilities) that was accessible from a Web services interface, and data
display though Web status pages.  Inca was first deployed to TeraGrid in mid
2003 and released in late 2003. After running Inca 1 for a year and a half on
TeraGrid, we learned some valuable lessons and designed Inca version 2, our
current release.  Inca 2 contains a number of substantial improvements over
Inca 1 with respect to security, installation and maintenance, and storage
capabilities.  A pre-release of Inca 2 was provided in February 2006 and a
beta release in November 2006.  A production version of Inca 2 was released in
January 2007.  Currently, seven other Grids in the U.S., Europe, and Australia
use Inca.

% This motivated our development of a new version of Inca.
In the next section, we describe our experiences with deploying Inca 1 to
TeraGrid and the lessons learned.  We then describe the Inca 2 architecture
and the benefits it provides.  Use cases illustrate that Inca 2 is being used
today and has great results.

\Section{Features}

@@

%------------------------------------------------------------------------- 
%------------------------------------------------------------------------- 
\Section{Architecture}

@@

\Section{Use Cases}

>>>>

% Complex means they are difficult to provide and maintain. Describe the
% stakeholders.   Monitoring is needed.
Providing and maintaining a stable infrastructure for these complex 
Grid systems poses challenges for both administrators who build and
maintain Grid resources and scientists who use them.  First, system
administration is distributed across multiple administrative domains which
requires a significant amount of coordination.  This is typically performed by
a group of Grid operators who take into account local site policies 
and decide when software (including updates and patches) should be deployed to
resources and how it should be configured for the set of heterogenous
resources.  Each site's system administrators, who are oftentimes not Grid
experts, are then responsible for providing the required Grid software
components and services for their resource and debugging any problems that
arise.  Well-defined requirements and good communication are required in this
model.  Otherwise inconsistencies between the resources arise that can either
inconvenience the user or prevent them from using particular resources
together.  Second, as a large distributed system, failures will occur over
time in the Grid services due to network or system failures, software
misconfiguration, or software bugs.  In order to detect such problems, 
monitoring of the resources is required.

% What types of Grid monitoring are around and what the limitations are
One approach to monitoring Grids, used by tools such as
MonALISA~\cite{monalisa} and GridICE~\cite{gridice}, is to aggregate and
display data from existing cluster or system administrator monitoring tools
such as Ganglia~\cite{ganglia}, CluMon~\cite{clumon}, and
Nagios~\cite{nagios}.  This provides a centralized, systems-level view of Grid
resources where low-level host statistics and queue information can be
examined.  This type of monitoring information is useful for showing the
utilization of Grid resources, however, it does not provide the type of
high-level monitoring needed to detect user problems within the Grid
infrastructure, such as incompatible software versions.

% What user-level monitoring provides and why we think it is good
User-level Grid monitoring approaches testing and performance measurement of
the Grid infrastructure from the user perspective.  The goal is to act as an
impartial test user of the Grid and detect problems with the infrastructure so
they can be fixed before users notice them -- user complaints should not be
the first indication of Grid failures.  We refer to an executable program that
tests or measures some aspect of a system or installed software from the
user-perspective as a reporter.  User-level Grid monitoring has the following
characteristics:

\begin{itemize}

\item In order to reflect regular user experiences, tests or performance
measurements of the Grid infrastructure should be done from a standard user
account.  In fact, it is important to not execute under a system
administrator's account because they may be privileged and often have custom
shell initialization files.

\item Also in order to reflect regular user experiences, tests should be
written and configured using information directly from user documentation
(e.g., hostnames, ports, pathnames, etc.).  This may not always be possible
when documentation and testing are done in parallel in a pre-production
environment but they should be closely coordinated activities.

\item Since Grids are dynamic, user-level tests or performance measurements
should be automated and executed periodically.

\item Tests or performance measurements that interact with Grid services
require authentication and should be executed using a standard user GSI
credential that is mapped to the standard user account.

\item Tests or performance measurements should be executed locally on all Grid
resources when appropriate so that all Grid access points available to users
are verified.  Similarly, it is important to executed some tests all-to-all
in order to detect site-to-site configuration errors such as authentication
problems.

\end{itemize}

\noindent In our experience, writing and maintaining user-level tests and
performance measurements is an iterative refinement process since Grids are
dynamic and the software environment changes over time as packages are
upgraded.  It is also difficult to write and deploy a test perfectly the first
time because of portability issues and sometimes incomplete user
documentation.  For example, it may take a few iterations to determine whether
detected failures are the result of a faulty test, fault user documentation,
or a faulty system.  Realistically, this is a labor-intensive activity and
thus the amount of user-level Grid monitoring is often limited by the amount
of people resources that can be assiged to it.  However, A good goal is to
test all the commands that exist in the user documentation and/or have at
least one basic test for each software component of the Grid infrastructure.

% Inca 1~\cite{inca1} first addressed user-level Grid monitoring but had
% limitations.  We learned lessons from our TeraGrid deployment.  
In 2003, SDSC, in partnership with TeraGrid, began developing Inca
1~\cite{inca1} to implement a user-level monitoring system with the above
features and started using it to validate and
verify that CTSS was deployed consistently across all TeraGrid resources and
to monitor its status.  At that time, the only available user-level Grid
monitoring tools were the NCSA TestGrid script~\cite{ncsa-test} and
GITS~\cite{gits}, scripts that run a fixed number of Grid tests and output
results in HTML.  Although these tools were easy to install and produced
useful information, they showed only the view of the Grid from a single
resource, lacked automation, and were not easily extensible.  The initial
version of Inca was implemented as a client-server architecture.  It provided
data collection, data storage (with limited archiving capabilities) that was
accessible from a Web services interface, and data display though Web status
pages.  Inca was first deployed to TeraGrid in mid 2003 and released in late
2003. After running Inca 1 for a year and a half on TeraGrid, we learned some
valuable lessons and designed Inca 2, our current release.  Inca 2
contains a number of substantial improvements over Inca 1 with respect to
security, installation and maintenance, and storage capabilities.  A
production release of Inca 2 was provided in February 2007 and has been
running on TeraGrid since November 2006.  Currently, seven other Grids in the
U.S., Europe, and Australia use Inca.

% This motivated our development of a new version of Inca.
In the next section, we describe the features of Inca 2 and note which
are new features based on our experiences with deploying Inca 1 on TeraGrid.
We then describe the Inca 2 architecture and how it can be used to provide
user-level Grid monitoring.  In Section~\ref{usecases}, we describe two 
uses of Inca 2 for Grid software environment verification and benchmarking.  
Finally, we describe some future work and summarize the paper.

%------------------------------------------------------------------------- 
\Section{Inca 2 Features}

@@

%------------------------------------------------------------------------- 
\Section{Architecture}

@@

\Section{Use Cases}
\label{usecases}
